### 2019-02-14 gpus/paper.tex

%\documentclass[12pt, preprint]{aastex}
\documentclass[twocolumn,times]{aastex62}
%\documentclass[preprint,times]{aastex6}

% these lines seem necessary for pdflatex to get the paper size right
\pdfpagewidth 8.5in
\pdfpageheight 11.0in

\usepackage{epsf,color,amsmath}

\usepackage{cancel}

\newcommand{\sfrac}[2]{\mathchoice%
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
    \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
    \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
    \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}} {#1\!/#2}}

\newcommand{\castro}{{\sf Castro}}
\newcommand{\maestro}{{\sf Maestro}}

\newcommand{\nablab}{{\mathbf{\nabla}}}
\newcommand{\Ub}{\mathbf{U}}
\newcommand{\gb}{\mathbf{g}}
\newcommand{\omegadot}{\dot{\omega}}
\newcommand{\Sdot}{\dot{S}}
\newcommand{\ddx}[1]{{\frac{{\partial#1}}{\partial x}}}
\newcommand{\ddt}[1]{{\frac{{\partial#1}}{\partial t}}}
\newcommand{\odt}[1]{{\frac{{d#1}}{dt}}}
\newcommand{\divg}[1]{{\nablab \cdot \left (#1\right)}}

\newcommand{\Ic}{\mathcal{I}}
\newcommand{\smax}{{s_\mathrm{max}}}

\usepackage{bm}

\newcommand{\Uc}{{\bm{\mathcal{U}}}}
\newcommand{\Fb}{\mathbf{F}}
\newcommand{\Sc}{\mathbf{S}}

\newcommand{\xv}{{(x)}}
\newcommand{\yv}{{(y)}}
\newcommand{\zv}{{(z)}}

\newcommand{\ex}{{\bf e}_x}
\newcommand{\ey}{{\bf e}_y}
\newcommand{\ez}{{\bf e}_z}

\newcommand{\Ab}{{\bf A}}
\newcommand{\Sq}{{\bf S}_\qb}
\newcommand{\Sqhydro}{{\Sq^{\mathrm{hydro}}}}
\newcommand{\qb}{{\bf q}}

\newcommand{\Shydro}{{{\bf S}^{\mathrm{hydro}}}}
\newcommand{\Rb}{{\bf R}}
\newcommand{\Rq}{{\bf R}}
\newcommand{\Adv}[1]{{\left [\mathcal{A} \left(#1\right)\right]}}
\newcommand{\Advt}[1]{{\left [\mathcal{\tilde{A}} \left(#1\right)\right]}}
\newcommand{\Advs}[1]{{\mathcal{A} \left(#1\right)}}

\setlength{\marginparwidth}{0.75in}
\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}

\begin{document}
%======================================================================
% Title
%======================================================================
\title{CASTRO: A New Compressible Astrophysical Solver. IV. Performance Portability}

\shorttitle{}
\shortauthors{}

\correspondingauthor{Michael Zingale}

\collaboration{the Castro Development Team}

\author[0000-0003-2103-312X]{Ann S. Almgren}
\affil{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory}

\author[0000-0002-3185-9809]{Maria Barrios Sazo}
\affil{Department of Physics and Astronomy, Stony Brook University}

\author{Kevin Gott}
\affil{National Energy Research Scientific Computing Center}

\author[0000-0002-1530-781X]{Alice Harpole}
\affil{Department of Physics and Astronomy, Stony Brook University}

\author[0000-0003-0439-4556]{Max P.~Katz}
\affil{NVIDIA Corporation}

\author[0000-0003-2300-5165]{Donald Willcox}
\affil{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory}

\author[0000-0001-8092-1974]{Weiqun Zhang}
\affil{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory}

\author[0000-0001-8401-030X]{Michael Zingale}
\affil{Department of Physics and Astronomy, Stony Brook University}
\email{michael.zingale@stonybrook.edu}

%======================================================================
% Abstract and Keywords
%======================================================================
\begin{abstract}
We describe a performance portable version of the \castro\ hydrodynamics
code that is able to achieve excellent scaling on both CPU and GPU-based
supercomputers.
\end{abstract}

\keywords{hydrodynamics---methods: numerical}

%======================================================================
% Introduction
%======================================================================
\section{Introduction}\label{Sec:Introduction}

\castro\ is a compressible (radiation-)hydrodynamics simulation code built on
the AMReX adaptive mesh refinement framework \citep{castro,castro2,castro3}.  It 
includes full self-gravity using multigrid methods with isolated boundaries
\citep{katz:2016}, works with general equations of state \citep{zingalekatz}
and arbitrary nuclear reaction networks.  Castro has been used for science investigations of
...

The original computing architectures targeted were

With GPU accelerators becoming increasingly dominant, algorithms need to be ported.
There are a number of descriptions of GPU-enabled astrophysical hydro codes, including
\citet{gamer,cholla,fargo3d,pekkila:2017,caplan:2018,racz:2018,liska:2018,goz:2018}

GPU-enabled integration has been explored by \citep{brock:2015}.

Discuss frameworks vs.\ doing it outselves

Discuss OpenMP, OpenACC, and CUDA.

In this paper, we describe our approach to performance portability.
Our goal is to have the same computational kernels run efficiently on
CPUs (including manycore processors) and GPUs.  We describe the work
we needed to do to enable this portability and show performance
metrics for a reacting flow problem in astrophysics.

\section{The Castro hydrodynamics code}

Disucss changes since last paper

\castro\ is developed using a fully open-development model, with the
code managed in git and hosted on
github\footnote{\url{https://github.com/amrex-astro/Castro/}}.
Contributions from the community are welcomed via github issues and
pull requests.  For the science projects that the core development
team are working on, all problem setup, input files, auxillary data,
and anything else needed for the science is stored in the main github
repo, allowing for the community to contribute to the science and
reproduce results as they are published.

\section{Performance portable compute kernels}

Describe mfiter structure

Describe MPI, MPI+OpenMP w/ tiling, MPI + CUDA

\begin{figure*}
\centering
\includegraphics[width=0.3\textwidth]{gpu_1} 
\includegraphics[width=0.3\textwidth]{gpu_2} 
\includegraphics[width=0.3\textwidth]{gpu_3}
\caption{\label{fig:loops} caption}
\end{figure*}

\subsection{Summary of changes required}

The main hydrodynamics solver in \castro\ is an unsplit corner
transport upwind \citep{ppmunsplit} piecewise parabolic method
\citep{ppm} scheme.  This uses full corner coupling, requiring a 
number of transverse Riemann problem solutions.

\section{Reaction networks}

\section{Scaling}

Our current focus is on offloading the hydrodynamics and reactions to
GPUs, so we will achieve the best GPU performance for problems without
self-gravity.

\subsection{Pure hydrodynamics}

\subsection{Reaction test}

\subsection{X-ray burst}

This problem uses constant gravity, so all of the physics can be done on the GPUs.

\section{Summary}

Future work: gravity on GPUs (multigrid or multipole), port of \maestro\ following
the same ideas.

\software{MPICH, GCC, Castro, python, matplotlib}

\acknowledgements \castro\ is freely available at
\url{http://github.com/AMReX-Astro/Castro}.  The work at Stony Brook
was supported by DOE/Office of Nuclear Physics grant DE-FG02-87ER40317
and NSF award AST-1211563.  An award of computer time was provided by
the Innovative and Novel Computational Impact on Theory and Experiment
(INCITE) program.  This research used resources of the Oak Ridge
Leadership Computing Facility at the Oak Ridge National Laboratory,
which is supported by the Office of Science of the U.S. Department of
Energy under Contract No. DE-AC05-00OR22725.  We thank NVIDIA Corporation
for the donation of a Titan X Pascal used in this research.

%======================================================================
% References
%======================================================================

\bibliographystyle{apj}
\bibliography{ws}

\end{document}