### 2019-03-18 MAESTROeX_I/conclusions.tex
more ferences

\section{Conclusions and Future Work}
Our new implementation of the MAESTRO code in the AMReX software framework

Our future scientific investigations include convection in massive rotating stars, \citep{heger2000presupernova}, the convective Urca process in white dwarfs \citep{willcox2016type}, solar physics \citep{wood2018self}, and magnetohydrodynamics \citep{wood2015three,wood2011sun}.

Our future algorithmic enhancements include more accurate and higher-order multiphysics coupling strategies byased on spectral deferred corrections \cite{dutt2000spectral,bourlioux2003high}.
This framework has been successfully used in terrestrial combustion \cite{pazner2016high,nonaka2018conservative}

Our software enhancements will include GPU implementation.
\MarginPar{Forward cite CASTRO paper; add a few sentences on GPU strategy}

\acknowledgements

The work at LBNL was supported by the U.S. Department of Energy's Scientific Discovery Through Advanced Computing (SciDAC) program under contract No. DE-AC02-05CH11231.
This research used resources of the National Energy Research Scientific Computing Center (NERSC), a U.S. Department of Energy Office of Science User Facility operated under Contract No. DE-AC02-05CH11231.